\Large \bf FlexAlloc: Dynamic Memory Partitioning\\ between the Corevisor and Hostvisor in SeKVM

### 2023-12-22 paper.tex
Add barebones evaluation

All of the source code described below can be found at \url{https://github.com/nelsonm2991/host-sekvm-project}.

@@

We evaluate FlexAlloc by TODO

@@

@@

Another key shortcoming of our project is that it does not fully implement
dynamic memory partitioning between the corevisor and hostvisor. As mentioned in
the introduction, another area in which corevisor memory scales with the number
of VMs is in maintaining \textit{struct el2\_vm\_info} objects that record the
metadata associated with each guest VM. Currently, we still statically allocate
a fixed-size array of \textit{struct el2\_vm\_info} objects in corevisor
memory ~\cite{hypsec_host.h}. This essentially preserves the hard cap on the
number of guest VMs that can be run. An easy fix would be to keep this array
as a statically-allocated array, while increasing the number of elements. Since each
\textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a 
significant amount of memory even if fewer than sixteen VMs are created.

We also do not fully evaluate our implementation by trying to run a large number
of VMs. To do this, we would have to modify the macros used to define the current
sixteen-VM limit. It is entirely conceivable that the hostvisor might not be able
to allocate 1M contiguous regions of memory past a certain point if memory
pressure and external fragmentation are high. In this case, the hostvisor would
not be able to allocate the memory regions needed for the corevisor to create
the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
the VM stage-2 page tables down into even smaller regions of memory, but this would
likely lead to more complicated logic when traversing the page table.

Another possible area of future work is exploring other allocators designed for
allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables
into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
to allocate contiguous regions larger than 4M. It might be possile to avoid this
by looking at other Linux memory allocators explicitly designed for the purpose
of large contiguous allocations.

>>>>

All of the source code described below can be found at \url{https://github.com/nelsonm2991/host-sekvm-project}.
Our implementation of FlexAlloc is based on SeKVM, which is in turn based on Linux v5.4.55.

@@

We evaluate FlexAlloc by running it on a QEMU-emulated Raspberry Pi 4. In
particular, we run VMWare Fusion Professional Version 13.0.2 on an M1 Macbook
Air running MacOS Ventura 13.6. Using VMWare, we create a VM
with 2 processor cores and 8704 MB memory running Ubuntu 22.04.3 LTS. On this
VM, we run QEMU 6.2.0 to emulate a Raspberry Pi 4. This emulated Raspberry
Pi 4 serves as our host machine, on which we run FlexAlloc. Our guest VMs 
run vanilla Linux v5.4.

Using FlexAlloc, we boot up a guest VM and verify that it is able to start up
without any kernel panics. We use \textit{print\_string} to verify that we
are passing eight distinct 1M regions to the corevisor when the VM is created.
We test a case where the 1M regions are not contiguous, and verify that our
modified page table iterators are able to successfully handle a VM stage-2 page
table that is non-contiguous. We also shut down the guest VM and verify that
all eight 1M regions are returned to the hostvisor. We compare the base address
of each region allocated during VM creation against the base address of each
region reclaimed during VM shutdown, and verify that they match each other.

@@

\subsection{Shortcomings in design}

@@

Another key shortcoming of our design is that it does not fully implement
dynamic memory partitioning between the corevisor and hostvisor. As mentioned in
the introduction, another area in which corevisor memory scales with the number
of VMs is in maintaining \textit{struct el2\_vm\_info} objects that record the
metadata associated with each guest VM. Currently, we still statically allocate
a fixed-size array of \textit{struct el2\_vm\_info} objects in corevisor
memory ~\cite{hypsec_host.h}. This essentially preserves the hard cap on the
number of guest VMs that can be run. An easy fix would be to keep this array
as a statically-allocated array, while increasing the number of elements. Since each
\textit{struct el2\_vm\_info} object is only 768 bytes, this would not waste a 
significant amount of memory even if fewer than sixteen VMs are created.

\subsection{Shortcomings in evaluation}

We also do not fully evaluate our implementation. For one, we do not stress-test our implementation by running a large number
of VMs. To do this, we would have to modify the macros used to define the current
sixteen-VM limit. It is entirely conceivable that the hostvisor might not be able
to allocate 1M contiguous regions of memory past a certain point if memory
pressure and external fragmentation are high. In this case, the hostvisor would
not be able to allocate the memory regions needed for the corevisor to create
the stage-2 page table for a new VM. We could perhaps delay this limit by breaking
the VM stage-2 page tables down into even smaller regions of memory, but this would
likely lead to more complicated logic when traversing the page table. In addition, we do not verify whether fragmenting the stage-2 page table results
in any degradation in VM performance. Ideally, we would run a memory-intensive
benchmark on our guest VMs to verify that our iterators work properly under a
high memory load. Finally, we do not quantify the change in corevisor memory usage.
We assume that by not allocating the memory for VM stage-2 page tables at bootup,
corevisor memory will automatically be reduced by an amount matching the
size of this static allocation. This should be verified using actual data. For
instance, we could periodically iterate over every page and see how many pages
are owned by the corevisor before and after VM creation.

\subsection{Future work}

A possible area of future work is exploring other allocators designed for
allocating large chunks of contiguous memory. We fragment the VM stage-2 page tables
into eight distinct 1M pages, because \textit{alloc\_pages} does not allow us
to allocate contiguous regions larger than 4M. It might be possile to avoid this
by looking at other Linux memory allocators explicitly designed for the purpose
of large contiguous allocations.
